---
title: "TDLMM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TDLMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

This vignette demonstrates the implementation of treed distributed lag mixture model (TDLMM). More details can be found in Mork and Wilson (2023) <doi: [10.1111/biom.13568](https://doi.org/10.1111/biom.13568)>.

```{r setup, message = FALSE}
library(dlmtree)
library(tidyverse)
set.seed(1)
data("sbd_dlmtree")
```

### Data preparation
```{r}
# Response and covariates
sbd_cov <- sbd_dlmtree %>% 
            select(bwgaz, ChildSex, MomAge, GestAge, MomPriorBMI, Race,
                    Hispanic, MomEdu, SmkAny, Marital, Income,
                    EstDateConcept, EstMonthConcept, EstYearConcept)

# Exposure data
sbd_exp <- list(PM25 = sbd_dlmtree %>% select(starts_with("pm25_")),
                TEMP = sbd_dlmtree %>% select(starts_with("temp_")),
                SO2 = sbd_dlmtree %>% select(starts_with("so2_")),
                CO = sbd_dlmtree %>% select(starts_with("co_")),
                NO2 = sbd_dlmtree %>% select(starts_with("no2_")))
sbd_exp <- sbd_exp %>% lapply(as.matrix)
```

### Fitting the model
```{r tdlmm.fit}
tdlmm.fit <- dlmtree(formula = bwgaz ~ ChildSex + MomAge + MomPriorBMI +
                                        Race + Hispanic + SmkAny + EstMonthConcept,
                     data = sbd_cov,
                     exposure.data = sbd_exp,
                     family = "gaussian", dlm.type = "linear", mixture = TRUE,
                     mixture.interactions = "noself",
                     n.burn = 2500, n.iter = 10000, n.thin = 5)
```

### Model fit summary
```{r tdlmm.sum}
# Marginalization with co-exposure fixed at exact levels for each exposure
tdlmm.sum.exact <- summary(tdlmm.fit, marginalize = c(3, 2, 1, 2, 1))

# Marginalization with co-exposure fixed at 25th percentile
tdlmm.sum.percentile <- summary(tdlmm.fit, marginalize = 25)

# Marginalization with co-exposure fixed at the empirical means (default)
tdlmm.sum <- summary(tdlmm.fit, marginalize = "mean", log10BF.crit = 0.5)
tdlmm.sum
```

### Main exposure effect
```{r tdlmm.plot}
library(gridExtra)

p1 <- plot(tdlmm.sum, exposure1 = "PM25", main = "PM2.5")
p2 <- plot(tdlmm.sum, exposure1 = "TEMP", main = "Temperature")
p3 <- plot(tdlmm.sum, exposure1 = "SO2", main = "SO2")

grid.arrange(p1, p2, p3, nrow = 1)
```

### Lagged interaction effect
```{r}
plot(tdlmm.sum, exposure1 = "PM25", exposure2 = "TEMP")
```

### Adjusting for expected changes in co-occurring exposures
Here we consider going from the 25th to the 75th percentile in each exposure while adjusting for the expected changes in other exposures due to their correlations with the exposure of interest.
```{r}
library(ggplot2)
dlm_coexp <- adj_coexposure(sbd_exp, tdlmm.fit, contrast_perc = c(0.25, 0.75))
ggplot(dlm_coexp, aes(x = Time, y = Effect, ymin = Lower, ymax = Upper)) +
  geom_hline(yintercept = 0, color = "red") +
  geom_ribbon(fill = "grey") +
  geom_line() +
  facet_wrap(~Name) +
  theme_bw() + 
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))
```