---
title: "TDLNM"
resource_files:
  - vignettes/articles/sbd_dlmtree.rda
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates the implementation of treed distributed lag non-linear model (TDLNM). More details can be found in Mork and Wilson (2021) <doi: [10.1093/biostatistics/kxaa051](https://doi.org/10.1093/biostatistics/kxaa051)>.

```{r setup, message = FALSE}
library(dlmtree)
library(dplyr)
set.seed(1)
```

### Load data

Simulated data is available on [GitHub](https://github.com/danielmork/dlmtree/tree/master/vignettes/articles). It can be loaded with the following code.

```{r load data from GitHub, eval=FALSE, include=TRUE}
sbd_dlmtree <- get_sbd_dlmtree()
```


```{r load data, include=FALSE}
load("sbd_dlmtree.rda")
```


### Data preparation
```{r}
# Response and covariates
sbd_cov <- sbd_dlmtree %>% 
            select(bwgaz, ChildSex, MomAge, GestAge, MomPriorBMI, Race,
                    Hispanic, MomEdu, SmkAny, Marital, Income,
                    EstDateConcept, EstMonthConcept, EstYearConcept)

# Exposure data
sbd_exp <- list(PM25 = sbd_dlmtree %>% select(starts_with("pm25_")),
                TEMP = sbd_dlmtree %>% select(starts_with("temp_")),
                SO2 = sbd_dlmtree %>% select(starts_with("so2_")),
                CO = sbd_dlmtree %>% select(starts_with("co_")),
                NO2 = sbd_dlmtree %>% select(starts_with("no2_")))
sbd_exp <- sbd_exp %>% lapply(as.matrix)
```

### Fitting the model
```{r tdlnm.fit}
tdlnm.fit <- dlmtree(formula = bwgaz ~ ChildSex + MomAge + MomPriorBMI + 
                       Race + Hispanic + SmkAny + EstMonthConcept,
                     data = sbd_cov,
                     exposure.data = sbd_exp[["TEMP"]],
                     dlm.type = "nonlinear",
                     family = "gaussian",
                     control.tdlnm = list(exposure.splits = 20),
                     control.mcmc = list(n.burn = 2500, n.iter = 10000, n.thin = 5))
```

### Model fit summary
```{r tdlnm.sum}
tdlnm.sum <- summary(tdlnm.fit)
print(tdlnm.sum)
```

### Exposure-time surface
```{r tdlnm.plot}
plot(tdlnm.sum, 
     main = "Plot title", 
     xlab = "Time axis label", 
     ylab = "Exposure-concentration axis label", 
     flab = "Effect color label")
```

### Slicing on exposure-concentration
```{r}
# slicing on exposure-concentration
plot(tdlnm.sum, plot.type = "slice", val = 1, main = "Slice at concentration 1") 
plot(tdlnm.sum, plot.type = "slice", val = 2, main = "Slice at concentration 2")
```

### Slicing on time lag
```{r}
# slicing on exposure-concentration
plot(tdlnm.sum, plot.type = "slice", time = 7, main = "Slice at time 7")
plot(tdlnm.sum, plot.type = "slice", time = 15, main = "Slice at time 15")
plot(tdlnm.sum, plot.type = "slice", time = 33, main = "Slice at time 33")
```

### different plot.type options
```{r}
# Standard error, credible intervals
plot(tdlnm.sum, plot.type = "se", main = "Standard error")  
plot(tdlnm.sum, plot.type = "ci-min", main = "Credible interval lower bound")
plot(tdlnm.sum, plot.type = "ci-max", main = "Credible interval upper bound")
```

```{r}
# Cumulative effect and significance
plot(tdlnm.sum, plot.type = "cumulative", main = "Cumulative effect per exposure-concentration")
plot(tdlnm.sum, plot.type = "effect", main = "Significant effects with directions")
```





